---
title: "DSMolgenisArmadillo"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Build Status](https://api.travis-ci.com/molgenis/molgenis-r-datashield.svg?branch=master)](https://app.travis-ci.com/github/molgenis/molgenis-r-datashield/)
[![CRAN status](https://www.r-pkg.org/badges/version/DSMolgenisArmadillo)](https://CRAN.R-project.org/package=DSMolgenisArmadillo)
[![codecov](https://codecov.io/gh/molgenis/molgenis-r-datashield/branch/master/graph/badge.svg?token=b661qHT3BO)](https://app.codecov.io/gh/molgenis/molgenis-r-datashield)
<!-- badges: end -->

A DSI implementation for the [MOLGENIS Armadillo DataSHIELD Service](https://github.com/molgenis/molgenis-service-armadillo/).

## Overview
DSMolgenisArmadillo is an extension to DSI which implements S4 classes and generic methods for Armadillo. It extends the functionality of DSI to work with Armadillo servers and is necessary for DataSHIELD analysis using Armadillo.

For more detailed documentation check: [https://github.com/datashield/DSI](https://github.com/datashield/DSI/).

## Installation
You can install the released version of DSMolgenisArmadillo from [CRAN](https://CRAN.R-project.org) with:

```{r eval = FALSE}
install.packages("DSI")
install.packages("DSMolgenisArmadillo")
```

## Documentation
Check the [package documentation](https://molgenis.github.io/molgenis-r-datashield/articles/DSMolgenisArmadillo.html) for details.

## Armadillo 2
The newest version (2.x) of DSMolgenisArmadillo will be only compatible with Armadillo version 3. If you still use Armadillo 2, you should use the 1.4.1 version of DSMolgenisArmadillo You can install this specific version the following commands:

For windows:
``` R
packageurl <- "https://cran.rstudio.com/bin/windows/contrib/4.2/DSMolgenisArmadillo_1.4.1.zip"
install.packages(packageurl, repos=NULL, type="source")
```

For Mac:
``` R
packageurl <- "hhttps://cran.rstudio.com/bin/macosx/contrib/4.2/DSMolgenisArmadillo_1.4.1.tgz"
install.packages(packageurl, repos=NULL, type="source")
```

For Linux:
``` R
packageurl <- "https://cran.rstudio.com/src/contrib/DSMolgenisArmadillo_1.4.1.tar.gz"
install.packages(packageurl, repos=NULL, type="source")
```

